---
title: "redist 101 - NC Example"
author: "Melissa Wu"
date: "2022-09-12"
output:
  html_document: default
  pdf_document: default
---

# Setup

First, we load the required packages for analysis. If the packages are not installed yet, be sure to do so beforehand.

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(alarmdata)
library(redist)
```

# Download Simulation Data

We can access data from all 50 states in the US using `alarmdata`. Here, we use North Carolina as an example: `alarm_census_vest()` downloads geographic, demographic, and election data, `alarm_50state_map` loads the `redist_map` object that contains precinct geometries and adjacencies, and `alarm_50state_plans` downloads the 5000 pre-simulated redistricting plans from the [50-State Redistricting Simulations](https://doi.org/10.48550/arXiv.2206.10763).

```{r}
data_nc <- alarm_census_vest("NC")
map_nc <- alarm_50state_map("NC")
plans_nc <- alarm_50state_plans("NC")
```

If you are curious how these plans were generated, the code is available on the [50-State GitHub repository](https://github.com/alarm-redist/fifty-states/tree/main/analyses/NC_cd_2020). 

Specifically in North Carolina, districts must:
1. be contiguous
2. have equal populations
3. be geographically compact
4. preserve county boundaries as much as possible
Each of these redistricting requirements, in addition to the 1965 Voting Rights Act, were closely followed when conducting the simulations.

The 50-State `redist_plans` objects include data for the 2020 enacted districting plans. However, analysts might want to compare other custom plans to the 50-State simulations. This can be done with `alarm_add_plan()`. The following example adds the 2010 enacted plan for North Carolina as a reference plan to the `plans_nc` object.
```{r}
# error - 2010 plan has 13 districts, 2020 plan has 14
# plans_nc %>% alarm_add_plan(plans_nc, ref_plan = map_nc$cd_2010, map_nc, name = "cd_2010")
```

We can plot some of these plans to visualize the geographical boundaries of the simulated districts.

```{r}
redist.plot.plans(plans_nc, draws=1:4, shp=map_nc)
```

We can use `comp_edge` to measure the compactness of entire plans, and we can use `comp_polsby` to measure the compactness of individual districts across plans. 

```{r}
redist.plot.hist(plans_nc, qty = comp_edge) +
  labs(x = "Fraction of Edges Kept", y = "Percentage of Plans") +
  theme_bw()
```


We can also compare distributions of a certain summary statistic across districts. The following code plots the distribution of democratic percentage of each district for all simulations.

```{r}
redist.plot.distr_qtys(plans_nc, qty = pre_20_dem_bid / (pre_20_dem_bid + pre_20_rep_tru), geom = "boxplot") +
  labs(y = "2020 Presidential Democratic Vote Percentage") +
  theme_bw()
```

