---
title: "redist 101 - redist workflow 01"
author: "Sho Miyazaki"
date: "2022-09-15"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(alarmdata)
library(redist)
```

# Overview of building data objects

## `redist_map` 
The first step one in a `redist` analysis is creating the `redist_map` object, which stores the basic parameters of the redistricting problem. 
Here, we use the `redist_map` objects that can be downloaded from `alarmdata` package.
```{r message=FALSE}
map_nc <- alarm_50state_map("NC")
print(map_nc)
```
As it shows above, `redist_map` (and shapefiles) looks like an dataframe with many columns, such as populations and geographical information. 

For example, the `geometry` column contains the geographic shapefile information, which can be mapped as `POLYGON` (or `MULTIPOLYGON`) shape of the geographical area. 

```{r}
head(map_nc$geometry)
```

## Shapefiles and Projections
If you want to use the other shapefile data for the redistricting analysis (other countries, etc.), you need to transform the available shapefiles (census data, etc.) into the `redist_map` objects using [`redist_map()` function](https://alarm-redist.org/redist/reference/redist_map.html). 

As there are multiple formats of [shapefiles](https://r-spatial.github.io/sf/), you may also need to convert and adjust them into a unified format. 
For example, projections must be in the same style. For more information, check pp.17-18 of [this presentation for the workshop](https://docs.google.com/presentation/d/1FlTfDv8sjhDxxU2KIOnKd1SHhvivm7WdD0bYdrNNgo0/edit#slide=id.g11d65f9e588_0_0). 

## Adjacency List
Since the legislative district plans are generally required to be contiguous, all the `redist` simulation algorithms operate on an *adjacency graph*, which is constructed from the actual precinct or county geography.

What is adjacency? -- It is the formalization of contiguity. 

Then, what does contiguity means? -- Two units are adjacent if they are contiguous geographically. 

Let’s check what the adjacency list looks like.
The `adj` column of `redist_map` is the adjacency list. 

```{r}
head(map_nc$adj)
```

As you can see, the adjacency list shows that the area (row-wise) is adjacent/contiguous to which areas. 

With the `redist_map` object, we can plot this adjacency list. 
```{r}
plot(map_nc, adj=T)
```

You may need to (fix the adjacency list manually)][https://christophertkenny.com/geomander/reference/add_edge.html], particularly when there are islands connected with bridge. 

## Adding reference plans 
??? 

## Further Pre-processing
Often, we want to only analyze a portion of a map, or hold some districts fixed while others are re-simulated. We may also want to implement a status-quo-type constraint that encourages simulated districts to be close to a reference plan. This can be accomplished by freezing the “cores” of each district.

All of these operations fall under the umbrella of map pre-processing, and `redist` is well-equipped to handle them.  

The map [pre-processing vignette](https://alarm-redist.org/redist/articles/map-preproc.html) contains more information and examples about these operations.


Running simulations
redist_smc, redist_mergesplit
Constraints
Purpose is to operationalize legal requirements and norms
Pop tol (hard constraint)
Soft constraints
redist_plans output, which we use to do stuff
